---
title: "Analysis for 'Can politicians’ answers to Voting Advice Applications be trusted?'"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    keep_md: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Preparations

Load packages (see package information at the very end of this document)

```{r message=FALSE, warning=FALSE}
library(rio)
library(dplyr)
library(psych)
library(lavaan)
library(semTools)
library(sjlabelled)

```

Read data file

```{r}
df2019 <- import("../data/dat2019.xlsx")

```

Select variables used in the analysis and make sure the variable names are correct

```{r}
VAA_LR_items<-c("h26","h27","h25","h28","y19")
VAA_LR_items %in% names(df2019)

VAA_GT_items<-c("h21","h22","h13","h29","h24","y25")
VAA_GT_items %in% names(df2019)

CS_LR_items<-c("C2b","C2g","C2h")
CS_LR_items %in% names(df2019)

CS_GT_items<-c("C2a","C2c","C2d","C2e","C2f","C2i","C2j")
CS_GT_items %in% names(df2019)

#LR Self-placement
CS_LR_SP<-c("C5a")
CS_LR_SP %in% names(df2019)

#LR imagined voter placement
CS_LR_IP<-c("C5c")
CS_LR_IP %in% names(df2019)

Party_item<-c("puolue")
Party_item %in% names(df2019)

#vector for all item names

all_items<-c(Party_item,
             VAA_LR_items,
             VAA_GT_items,
             CS_LR_items,
             CS_GT_items,
             CS_LR_SP,
             CS_LR_IP)

#vector for all numeric item names

num_items<-c(
             VAA_LR_items,
             VAA_GT_items,
             CS_LR_items,
             CS_GT_items,
             CS_LR_SP,
             CS_LR_IP)

#vector for observed variables in CFA (and party)

obs_items<-c(Party_item,
             VAA_LR_items,
             VAA_GT_items,
             CS_LR_items,
             CS_GT_items)

#a vector for indicator variables

ind_items<-c(
             VAA_LR_items,
             VAA_GT_items,
             CS_LR_items,
             CS_GT_items)

#code all but party_item as numeric

df2019[,num_items]<-sapply(df2019[,num_items],as.numeric)
```

Print the responses to the observed items

```{r}
for (i in 1:length(obs_items)){
  print(obs_items[i])
  print(table(df2019[,obs_items[i]],useNA="always"))
  }
```

Data looks as it should!

Exclude completely missing cases

```{r}

df2019$completely_missing<-
  rowSums(is.na(df2019[,ind_items]))==length(ind_items)

#number of completely missing cases
table(df2019$completely_missing)

#proportion of completely missing cases
100*table(df2019$completely_missing)/nrow(df2019)

#filter the used sample
dat2019<-df2019 %>%
  filter(!completely_missing)


```

Transform/Reverse code high scores on observed variable to indicate right and TAN positioning

```{r}
reverse_items<-c("h26","y19",
                 "h21","h22","h13",
                 "C2g","C2h",
                 "C2c","C2e","C2i","C2j")

reverse_items %in% names(dat2019)

for (i in 1:length(reverse_items)){
  dat2019[,reverse_items[i]]<-6-dat2019[,reverse_items[i]]
}

```

\newpage

# Analysis

## Descriptive statistics

```{r}
#look what parties there are
cbind(n=table(dat2019$puolue),
      proportion=round(100*prop.table(table(dat2019$puolue)),2))

#how many responded to VAAs (any)

table(rowSums(is.na(dat2019[,c(VAA_LR_items,VAA_GT_items)]))!=
        length(c(VAA_LR_items,VAA_GT_items)))

#how many responded to CS (any)

table(rowSums(is.na(dat2019[,c(CS_LR_items,CS_GT_items)]))!=
        length(c(CS_LR_items,CS_GT_items)))

#table for CS-items

CS.item.table<-cbind.data.frame(
  item=c(CS_LR_items,CS_GT_items),
  description=c("The state should not interfere in economic activities",
                "Providing a stable social security network should be a state priority (r.)",
                "The state should take measures to reduce income disparities (r.)",
                "Immigrants should adapt to Finnish habits",
                "Stronger measures should be taken to protect the environment (r.)",
                "Same Sex Marriages should be prohibited by law",
                "Women should be favored in job search and promotion (r.)",
                "People who break the law should be punished more severely",
                "Immigrants are good for the Finnish economy (r.)",
                "Deciding on abortion issues should be a women's right (r.)"),
  round(data.frame(describe(dat2019[,c(CS_LR_items,CS_GT_items)],fast=T))[,c("n","mean","sd")],2))

export(CS.item.table,"../results/Table2_CS.items.xlsx",overwrite=T)

#table for VAA-items

VAA.item.table<-cbind.data.frame(
  item=c(VAA_LR_items,VAA_GT_items),
  description=c("If there will be a situation where one is forced to either cut public services and social benefits or increase taxes, tax increases are a better choice (r.)",
                "Large income inequalities are acceptable for compensating differences in people's talents and work ethic",
                "Public services should be outsourced more than they are now for private companies",
                "In the long run, the current extent of services and social benefits are too heavy for public economy",
                "Public authorities should be the main provider of social and healthcare services (r.)",
                "Gay and lesbian couples should have the same marriage and adoption rights as straight couples (r.)",
                "If the government proposes to establish a refugee center in my home municipality, the proposal should be accepted (r.)",
                "For Finland, the advantages of the EU outweigh the disadvantages (r.)",
                "Economic growth and creation of jobs should be given primacy over environmental issues, when these two collide",
                "Traditional values such as home, religion and fatherland form a good value base for politics",
                "Finland must adopt tough measures to defend order and protect regular citizens"),
  round(data.frame(describe(dat2019[,c(VAA_LR_items,VAA_GT_items)],fast=T))[,c("n","mean","sd")],2))

export(VAA.item.table,"../results/Table1_VAA.items.xlsx",overwrite=T)

```

\newpage

## H1 and H2

H1. Left-Right placement as computed from responses to the pre-election public Voting Advice Applications (VAAs) is positively associated with Left-Right placement as computed from responses to the privately administered post-election Candidate Survey (CS). This association is stronger than any associations between the Left-Right and GAL-TAN dimensions.

H2. GAL-TAN placement as computed from responses to the pre-election public Voting Advice Applications (VAAs) is positively associated with GAL-TAN placement as computed from responses to the privately administered post-election Candidate Survey (CS). This association is stronger than any associations between the Left-Right and GAL-TAN dimensions.

### Model script

```{r}


model_H1H2<-"
#loadings
VAA_LR=~h26+h27+h25+h28+y19
VAA_GT=~h21+h22+h13+h29+h24+y25
CS_LR=~C2b+C2g+C2h
CS_GT=~C2a+C2c+C2d+C2e+C2f+C2i+C2j

#latent correlations

#cross-dimension same-method
VAA_LR~~r.VAA*VAA_GT
CS_LR~~r.CS*CS_GT

#concurrent validity
VAA_LR~~r.LR*CS_LR
VAA_GT~~r.GT*CS_GT

#cross-dimension cross-method correlations
VAA_LR~~r.d1*CS_GT
VAA_GT~~r.d2*CS_LR

#custom parameters
test.H1:=r.LR-max(r.VAA,r.CS,r.d1,r.d2)
test.H2:=r.GT-max(r.VAA,r.CS,r.d1,r.d2)

"

```

### Fitting the model

```{r}
fit_H1H2<-cfa(model=model_H1H2,
              data=dat2019,
              missing="fiml")

```

Some problems with latent variable covariance structure

```{r}
lavInspect(fit_H1H2, "cov.lv")
lavInspect(fit_H1H2, "cor.lv")


#examine standardized estimates
std.est_H1H2<-standardizedsolution(fit_H1H2)
std.est_H1H2[std.est_H1H2$op=="~~" & 
               std.est_H1H2$lhs!=std.est_H1H2$rhs,]



```

There is an impossible correlation between GAL-TAN latent variables (absolute value > 1)

\newpage

### Respecified model: introduce the three preregistered residual correlations

Add the terms to the model script

```{r}
model_H1H2.re<-paste0(model_H1H2,
                      "h27~~C2h\n",
                      "h21~~C2d\n",
                      "h29~~C2c\n")
```

### Fitting the respecified model

```{r}
fit_H1H2.re<-cfa(model=model_H1H2.re,
              data=dat2019,
              missing="fiml")

```

### Results

Inspect fit of the model (first is the original model with problems,
second is the respecified)

```{r}
round(inspect(fit_H1H2,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)

round(inspect(fit_H1H2.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is adequate.

Hypotheses 1 and 2

Print standardized estimates to test the difference between correlations

```{r}
std.est_H1H2<-standardizedsolution(fit_H1H2.re)
std.est_H1H2[std.est_H1H2$op==":=" | 
               std.est_H1H2$op=="~~" & 
               std.est_H1H2$lhs!=std.est_H1H2$rhs,]

#save to a file
export(std.est_H1H2[std.est_H1H2$op!="~1",c(1:8)],
           "../results/Table3_Overall_H1H2_standardized_estimates.xlsx",
       overwrite=T)
```

H1: There is very strong (.915, p < .001) correlation between VAA-LR and CS-LR, and it is notably stronger (difference in correlations .492, p < .001) than the strongest of correlations between different dimensions (.424 between VAA_LR and VAA_GT, p < .001)

H2: There is very strong (.990, p < .001) correlation between VAA-GT and CS-GT, and it is notably stronger (difference in correlations .566, p < .001) than the strongest of correlations between different dimensions (.424 between VAA_LR and VAA_GT, p < .001)

\newpage

### Exploratory analysis for H1 and H2: Seek misspecification to improve the overall model fit

Residual correlations

```{r}
mis.rescor_H1H2<-miPowerFit(fit_H1H2.re,cor=.20)
mis.rescor_H1H2<-mis.rescor_H1H2[mis.rescor_H1H2$op=="~~" & 
                                   mis.rescor_H1H2$lhs!=mis.rescor_H1H2$rhs,]
#see summary of the decisions
table(mis.rescor_H1H2$decision.pow)

#there are 1 residual correlation that is a misspecification

rounded.vars<-c("mi","epc","target.epc",
                "std.epc","se.epc")

num.round<-function(var){
  var<-as.numeric(var)
  var<-round(var,2)
  return(var)
}

mis.rescor_H1H2[,rounded.vars]<-
  sapply(mis.rescor_H1H2[,rounded.vars],num.round)

printed.vars<-c("lhs","op","rhs","mi","epc","target.epc",
                "std.epc","std.target.epc","significant.mi",
                "high.power","decision.pow","se.epc")

#print the output

mis.rescor_H1H2 %>%
  filter(mis.rescor_H1H2$decision.pow=="M" | 
                mis.rescor_H1H2$decision.pow=="EPC:M") %>%
  dplyr::select(all_of(printed.vars)) 

  
```

There was one misspecified residual correlation in VAA-LR, between
h25. Public services should be outsourced more than they are now for private companies and y19. Public authorities should be the main provider of social and healthcare services (r.) 

### Exploratory respecification of the model

Add new parameter to the model script

```{r}
model_H1H2.exp.re<-paste0(model_H1H2.re,
                      "h25~~y19")

```

### Fitting the exploratory model

```{r}
fit_H1H2.exp.re<-cfa(model=model_H1H2.exp.re,
              data=dat2019,
              missing="fiml")
```

### Results from the exploratory model

```{r}
round(inspect(fit_H1H2.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)

round(inspect(fit_H1H2.exp.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is improved by additional residual correlation.

Retest Hypotheses 1 and 2

Print standardized estimates to test the difference between correlations

```{r}
std.est_H1H2.exp<-standardizedsolution(fit_H1H2.exp.re)
std.est_H1H2.exp[std.est_H1H2.exp$op==":=" | 
               std.est_H1H2.exp$op=="~~" & 
               std.est_H1H2.exp$lhs!=std.est_H1H2.exp$rhs,]

#save to a file
export(std.est_H1H2.exp[std.est_H1H2.exp$op!="~1",1:8],
           "../results/TableS1_Overall_H1H2.exp_standardized_estimates.xlsx",
       overwrite=T)
```

The results are virtually identical to those without the additional residual correlation. 

H1.exp: There is very strong (.932, p < .001) correlation between VAA-LR and CS-LR, and it is notably stronger (difference in correlations .462, p < .001) than the strongest of correlations between different dimensions (.470 between VAA_LR and VAA_GT, p < .001)

H2.exp: There is very strong (.990, p < .001) correlation between VAA-GT and CS-GT, and it is notably stronger (difference in correlations .520, p < .001) than the strongest of correlations between different dimensions (.470 between VAA_LR and VAA_GT, p < .001)

\newpage

## H5

H5. Left-Right self-placement in the privately administered post-election Candidate Survey (CS) is positively associated with Left-Right as computed from responses to the pre-election public Voting Advice Applications (VAAs). This association is stronger than the association between placement of an imagined party voter in the privately administered post-election Candidate Survey (CS) and Left-Right as computed from responses to the pre-election public Voting Advice Applications (VAAs).

### Add placement variables and their correlations with latent factors to the model used for H1 and H2

```{r}


model_H5<-"
#loadings
VAA_LR=~h26+h27+h25+h28+y19
VAA_GT=~h21+h22+h13+h29+h24+y25
CS_LR=~C2b+C2g+C2h
CS_GT=~C2a+C2c+C2d+C2e+C2f+C2i+C2j

#latent correlations

#cross-dimension same-method
VAA_LR~~r.VAA*VAA_GT
CS_LR~~r.CS*CS_GT

#convergent validity
VAA_LR~~r.LR*CS_LR
VAA_GT~~r.GT*CS_GT

#cross-dimension cross-method correlations
VAA_LR~~r.d1*CS_GT
VAA_GT~~r.d2*CS_LR

#residual correlations
h27~~C2h
h21~~C2d
h29~~C2c

#placement variables (defined as quasi-latent variables)

SP_LR=~C5a
IP_LR=~C5c

VAA_LR~~r.self.LR*SP_LR
VAA_LR~~r.ideal.LR*IP_LR

#custom parameters
test.H1:=r.LR-max(r.VAA,r.CS,r.d1,r.d2)
test.H2:=r.GT-max(r.VAA,r.CS,r.d1,r.d2)
test.H5:=r.self.LR-r.ideal.LR
"

```

### Fit the model

```{r}
fit_H5<-cfa(model=model_H5,
            data=dat2019,
            missing="fiml")
```


### Results

Inspect fit of the model

```{r}
round(inspect(fit_H5,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is adequate.

Hypothesis 5

Print standardized estimates to test the difference between correlations

```{r}
std.est_H5<-standardizedsolution(fit_H5)
std.est_H5[std.est_H5$op==":=" | 
               std.est_H5$op=="~~" & 
               std.est_H5$lhs!=std.est_H5$rhs,]

#save to a file
export(std.est_H5[std.est_H5$op!="~1",1:8],
           "../results/Table4_Overall_H5_standardized_estimates.xlsx",
       overwrite=T)
```

H5. The correlation between VAA_LR and CS Self-placement on LR is large (.829, p < .001) and larger than the association between VAA_LR and placement of imagined party voter (.739, p < .001; difference .090, p < .001)

### Exploratory H5: Seek misspecifications

Residual correlations

```{r}
mis.rescor_H5<-miPowerFit(fit_H5,cor=.20)
mis.rescor_H5<-mis.rescor_H5[mis.rescor_H5$op=="~~" & 
                                   mis.rescor_H5$lhs!=mis.rescor_H5$rhs,]
#see summary of the decisions
table(mis.rescor_H5$decision.pow)

#there are 1 residual correlation that is a misspecification

rounded.vars<-c("mi","epc","target.epc",
                "std.epc","se.epc")

num.round<-function(var){
  var<-as.numeric(var)
  var<-round(var,2)
  return(var)
}

mis.rescor_H5[,rounded.vars]<-sapply(mis.rescor_H5[,rounded.vars],num.round)

printed.vars<-c("lhs","op","rhs","mi","epc","target.epc",
                "std.epc","std.target.epc","significant.mi",
                "high.power","decision.pow","se.epc")

#print the output

mis.rescor_H5 %>%
  filter(mis.rescor_H5$decision.pow=="M" | 
                mis.rescor_H5$decision.pow=="EPC:M") %>%
  dplyr::select(all_of(printed.vars)) 

  
```

There was one misspecified residual correlation in VAA-LR, between
h25. Public services should be outsourced more than they are now for private companies and y19. Public authorities should be the main provider of social and healthcare services (r.) 


### Exploratory respecification of the model

```{r}
model_H5.exp<-paste0(model_H5,
                      "h25~~y19")

```

```{r}
fit_H5.exp<-cfa(model=model_H5.exp,
              data=dat2019,
              missing="fiml")
```

Inspect fit of the model

```{r}
round(inspect(fit_H5,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)

round(inspect(fit_H5.exp,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is improved.

Retest Hypothesis 5

Print standardized estimates to test the difference between correlations

```{r}
std.est_H5.exp<-standardizedsolution(fit_H5.exp)
std.est_H5.exp[std.est_H5.exp$op==":=" | 
               std.est_H5.exp$op=="~~" & 
               std.est_H5.exp$lhs!=std.est_H5.exp$rhs,]

#save to a file
export(std.est_H5.exp[std.est_H5.exp$op!="~1",1:8],
           "../results/TableS2_Overall_H5.exp_standardized_estimates.xlsx",
       overwrite=T)
```

The results are virtually identical to those without the additional residual correlation.

H5.exp. The correlation between VAA_LR and CS Self-placement on LR is large (.845, p < .001) and larger than the association between VAA_LR and placement of imagined party voter (.751, p < .001; difference .094, p < .001)


\newpage

## H3 and H4

H3. Within-party placement on Left-Right as computed from responses to the pre-election public Voting Advice Applications (VAAs) is positively associated with within-party placement on Left-Right as computed from responses to the privately administered post-election Candidate Survey (CS). This association is stronger than any within-party associations between the Left-Right and GAL-TAN dimensions.

H4. Within-party placement on GAL-TAN as computed from responses to the pre-election public Voting Advice Applications (VAAs) is positively associated with within-party placement on GAL-TAN as computed from responses to the privately administered post-election Candidate Survey (CS). This association is stronger than any within-party associations between the Left-Right and GAL-TAN dimensions. 

Construct a new dataframe that exclude other than members of the eight parties that have multiple members in the parliament

```{r}

dat2019.party<-dat2019 %>%
  filter(puolue=="KD" |
           puolue=="KESK" |
           puolue=="KOK" |
           puolue=="PS" |
           puolue=="RKP" |
           puolue=="SDP" |
           puolue=="VAS" |
           puolue=="VIHR")

table(dat2019.party$puolue)
```


### Model script

Add names for group specific parameters

```{r}


model_H3H4<-"
#loadings
VAA_LR=~h26+h27+h25+h28+y19
VAA_GT=~h21+h22+h13+h29+h24+y25
CS_LR=~C2b+C2g+C2h
CS_GT=~C2a+C2c+C2d+C2e+C2f+C2i+C2j

#latent correlations

#cross-dimension same-method
VAA_LR~~c(r.VAA.KD,r.VAA.KESK,r.VAA.KOK,r.VAA.PS,r.VAA.RKP,r.VAA.SDP,r.VAA.VAS,r.VAA.VIHR)*VAA_GT
CS_LR~~c(r.CS.KD,r.CS.KESK,r.CS.KOK,r.CS.PS,r.CS.RKP,r.CS.SDP,r.CS.VAS,r.CS.VIHR)*CS_GT

#convergent validity
VAA_LR~~c(r.LR.KD,r.LR.KESK,r.LR.KOK,r.LR.PS,r.LR.RKP,r.LR.SDP,r.LR.VAS,r.LR.VIHR)*CS_LR
VAA_GT~~c(r.GT.KD,r.GT.KESK,r.GT.KOK,r.GT.PS,r.GT.RKP,r.GT.SDP,r.GT.VAS,r.GT.VIHR)*CS_GT

#cross-dimension cross-method correlations
VAA_LR~~c(r.d1.KD,r.d1.KESK,r.d1.KOK,r.d1.PS,r.d1.RKP,r.d1.SDP,r.d1.VAS,r.d1.VIHR)*CS_GT
VAA_GT~~c(r.d2.KD,r.d2.KESK,r.d2.KOK,r.d2.PS,r.d2.RKP,r.d2.SDP,r.d2.VAS,r.d2.VIHR)*CS_LR

#custom parameters
mean.r.VAA:=mean(r.VAA.KD,r.VAA.KESK,r.VAA.KOK,r.VAA.PS,r.VAA.RKP,r.VAA.SDP,r.VAA.VAS,r.VAA.VIHR)
mean.r.CS:=mean(r.CS.KD,r.CS.KESK,r.CS.KOK,r.CS.PS,r.CS.RKP,r.CS.SDP,r.CS.VAS,r.CS.VIHR)
mean.r.LR:=mean(r.LR.KD,r.LR.KESK,r.LR.KOK,r.LR.PS,r.LR.RKP,r.LR.SDP,r.LR.VAS,r.LR.VIHR)
mean.r.GT:=mean(r.GT.KD,r.GT.KESK,r.GT.KOK,r.GT.PS,r.GT.RKP,r.GT.SDP,r.GT.VAS,r.GT.VIHR)
mean.r.d1:=mean(r.d1.KD,r.d1.KESK,r.d1.KOK,r.d1.PS,r.d1.RKP,r.d1.SDP,r.d1.VAS,r.d1.VIHR)
mean.r.d2:=mean(r.d2.KD,r.d2.KESK,r.d2.KOK,r.d2.PS,r.d2.RKP,r.d2.SDP,r.d2.VAS,r.d2.VIHR)

test.H3:=mean.r.LR-max(mean.r.VAA,mean.r.CS,mean.r.d1,mean.r.d2)
test.H4:=mean.r.GT-max(mean.r.VAA,mean.r.CS,mean.r.d1,mean.r.d2)

"

```


### Fit the configural model

```{r}
fit_H3H4<-cfa(model=model_H3H4,
              data=dat2019.party,
              group=c("puolue"),
              group.label=c("KD","KESK","KOK","PS","RKP","SDP","VAS","VIHR"),
              missing="fiml")

```

Problems with finding a converging model. Add preregistered residual correlations.

### Respecify model by adding the residual correlations

```{r}
model_H3H4.re<-paste0(model_H3H4,
                      "h27~~C2h\n",
                      "h21~~C2d\n",
                      "h29~~C2c\n")
```

### Fit the respecified model

```{r}
fit_H3H4.re<-cfa(model=model_H3H4.re,
              data=dat2019.party,
              group=c("puolue"),
              group.label=c("KD","KESK","KOK","PS","RKP","SDP","VAS","VIHR"),
              missing="fiml")

```

The problem persists

\newpage

## H3 and H4 with group-mean centered variables and no grouping structure

### ICC: Estimate how much of the variation in each item is between-groups

```{r}
#there was problems running the mult.icc function to the data structure so 
#data observed data was extracted from one of the previously fitted models
#to get rid of all labels etc.
num.dat.2019<-data.frame(fit_H5@Data@X,dat2019$puolue)
names(num.dat.2019)<-c(fit_H5@Data@ov$name,"puolue")
num.dat.2019<-num.dat.2019 %>%
  filter(puolue=="KD" |
           puolue=="KESK" |
           puolue=="KOK" |
           puolue=="PS" |
           puolue=="RKP" |
           puolue=="SDP" |
           puolue=="VAS" |
           puolue=="VIHR")


ICC<-data.frame(
  multilevel::mult.icc(x=num.dat.2019[,
                                      all_items[2:length(all_items)]],
                       grpid=num.dat.2019$puolue))
ICC[,2:3]<-round(ICC[,2:3],3)
ICC

describe(100*ICC$ICC1,fast=T)

#export 
export(ICC,"../results/TableS3_ICC.xlsx",overwrite=T)

```

ICC1 gives the proportion (%) of variance that is between the parties (ICC2 is the reliability of the group means). There is quite a lot of between-party variance, but the responses are not entire defined by party either.

\newpage

### Variable centering

```{r}

dat2019.gmc<-data.frame(dat2019.party)

na.mean<-function(var){
  mean(var,na.rm=T)
}

group.means<-dat2019.gmc %>%
  group_by(puolue) %>%
  summarise_at(all_items[2:length(all_items)],na.mean)


dat2019.gmc<-left_join(x=dat2019.gmc,
                       y=group.means,
                       by=c("puolue"),
                       suffix=c("",".pm"))



for(i in all_items[2:length(all_items)]){
  dat2019.gmc[i]<-
    dat2019.gmc[,i]-dat2019.gmc[,which(grepl(i,names(dat2019.gmc)) &
                           grepl("pm",names(dat2019.gmc)) & 
                     !grepl("r",names(dat2019.gmc))) ]
}

describe(dat2019.gmc[,all_items],fast=T)
```

\newpage

### Define the model

Identical to the model for H1 and H2

```{r}


model_H3H4<-"
#loadings
VAA_LR=~h26+h27+h25+h28+y19
VAA_GT=~h21+h22+h13+h29+h24+y25
CS_LR=~C2b+C2g+C2h
CS_GT=~C2a+C2c+C2d+C2e+C2f+C2i+C2j

#latent correlations

#cross-dimension same-method
VAA_LR~~r.VAA*VAA_GT
CS_LR~~r.CS*CS_GT

#concurrent validity
VAA_LR~~r.LR*CS_LR
VAA_GT~~r.GT*CS_GT

#cross-dimension cross-method correlations
VAA_LR~~r.d1*CS_GT
VAA_GT~~r.d2*CS_LR

#custom parameters
test.H3:=r.LR-max(r.VAA,r.CS,r.d1,r.d2)
test.H4:=r.GT-max(r.VAA,r.CS,r.d1,r.d2)

"

```



### Fit the model

```{r}
fit_H3H4<-cfa(model=model_H3H4,
              data=dat2019.gmc,
              missing="fiml")

```

Problems with latent variable covariance matrix

```{r}
lavInspect(fit_H3H4, "cov.lv")
lavInspect(fit_H3H4, "cor.lv")


```

There is a Heywood correlation between GAL-TAN latent variables (absolute value > 1)

\newpage

### Respecified model: introduce the three preregistered residual correlations

```{r}
model_H3H4.re<-paste0(model_H3H4,
                      "h27~~C2h\n",
                      "h21~~C2d\n",
                      "h29~~C2c\n")
```

### Fitting the respecified model

```{r}
fit_H3H4.re<-cfa(model=model_H3H4.re,
              data=dat2019.gmc,
              missing="fiml")

```

### Results

Inspect fit of the model

```{r}
round(inspect(fit_H3H4.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is quite poor according to CFI and TLI, but ok according to RMSEA and SRMR.

Hypotheses 3 and 4

Print standardized estimates to test the difference between correlations

```{r}
std.est_H3H4.re<-standardizedsolution(fit_H3H4.re)
std.est_H3H4.re[std.est_H3H4.re$op==":=" | 
               std.est_H3H4.re$op=="~~" & 
               std.est_H3H4.re$lhs!=std.est_H3H4.re$rhs,]

#save to a file
export(std.est_H3H4.re[std.est_H3H4.re$op!="~1",c(1:8)],
           "../results/Table3_Unconfounded_H3H4_standardized_estimates.xlsx",
       overwrite=T)
```

H3: There is strong (.782, p < .001) correlation between VAA-LR and CS-LR, and it is notably stronger (difference in correlations .593, p < .001) than the strongest of correlations between different dimensions (.189 between VAA_LR and VAA_GT, p = .006)

H4: There is very strong (.956, p < .001) correlation between VAA-GT and CS-GT, and it is notably stronger (difference in correlations .767, p < .001) than the strongest of correlations between different dimensions (.189 between VAA_LR and VAA_GT, p = .006)

\newpage

### Exploratory analysis for H3 and H4: Seek misspecification to improve the overall model fit

Residual correlations

```{r}
mis.rescor_H3H4<-miPowerFit(fit_H3H4.re,cor=.20)
mis.rescor_H3H4<-mis.rescor_H3H4[mis.rescor_H3H4$op=="~~" & 
                                   mis.rescor_H3H4$lhs!=mis.rescor_H3H4$rhs,]
#see summary of the decisions
table(mis.rescor_H3H4$decision.pow)

#there are 2 residual correlation that are misspecifications

rounded.vars<-c("mi","epc","target.epc",
                "std.epc","se.epc")

num.round<-function(var){
  var<-as.numeric(var)
  var<-round(var,2)
  return(var)
}

mis.rescor_H3H4[,rounded.vars]<-sapply(mis.rescor_H3H4[,rounded.vars],num.round)

printed.vars<-c("lhs","op","rhs","mi","epc","target.epc",
                "std.epc","std.target.epc","significant.mi",
                "high.power","decision.pow","se.epc")

#print the output

mis.rescor_H3H4 %>%
  filter(mis.rescor_H3H4$decision.pow=="M" | 
                mis.rescor_H3H4$decision.pow=="EPC:M") %>%
  dplyr::select(all_of(printed.vars)) 

  
```

There were two misspecified residual correlation.

One was between VAA-LR items (same misspecification as was found for H1 and H2)
H25. Public services should be outsourced more than they are now for private companies and y19. Public authorities should be the main provider of social and healthcare services (r.) 

The other misspecification was between C2a. Immigrants should adapt to Finnish habits and C2f. People who break the law should be punished more severely 

Re-specify the model to allow these parameters to be freely estimated

### Exploratory respecification of the model

```{r}
model_H3H4.exp.re<-paste0(model_H3H4.re,
                      "h25~~y19\n",
                      "C2a~~C2f\n")

```

### Fitting the exploratory model

```{r}
fit_H3H4.exp.re<-cfa(model=model_H3H4.exp.re,
              data=dat2019.gmc,
              missing="fiml")
```

### Results from the exploratory model

```{r}
round(inspect(fit_H3H4.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)

round(inspect(fit_H3H4.exp.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is improved

Retest Hypotheses 4 and 5

Print standardized estimates to test the difference between correlations

```{r}
std.est_H3H4.exp.re<-standardizedsolution(fit_H3H4.exp.re)
std.est_H3H4.exp.re[std.est_H3H4.exp.re$op==":=" | 
               std.est_H3H4.exp.re$op=="~~" & 
               std.est_H3H4.exp.re$lhs!=std.est_H3H4.exp.re$rhs,]


#save to a file
export(std.est_H3H4.exp.re[std.est_H3H4.exp.re$op!="~1",c(1:8)],
           "../results/TableS1_Unconfounded_H3H4.exp_standardized_estimates.xlsx",
       overwrite=T)

```

The results are virtually identical to those without the additional residual correlations.

H3: There is a strong (.831, p < .001) correlation between VAA-LR and CS-LR, and it is notably stronger (difference in correlations .592, p < .001) than the strongest of correlations between different dimensions (.239 between VAA_LR and VAA_GT, p < .001)

H4: There is a very strong (.957, p < .001) correlation between VAA-GT and CS-GT, and it is notably stronger (difference in correlations .718, p < .001) than the strongest of correlations between different dimensions (.239 between VAA_LR and VAA_GT, p < .001)

\newpage

## H6 with group mean centered observed variables

H6. Within-party placement on Left-Right as computed from responses to the pre-election public Voting Advice Applications (VAAs) is positively associated with within-party placement on Left-Right as computed from responses to the privately administered post-election Candidate Survey (CS). This association is stronger than any within-party associations between the Left-Right and GAL-TAN dimensions. 

### Add placement variables and their correlations with latent factors to the model used for H3 and H4

Model already includes the three preregistered correlations

```{r}

model_H6<-paste0(model_H3H4.re,
                 "SP_LR=~C5a\n",
                 "IP_LR=~C5c\n",
                 "VAA_LR~~r.self.LR*SP_LR\n",
                 "VAA_LR~~r.ideal.LR*IP_LR\n",
                 "test.H6:=r.self.LR-r.ideal.LR\n")


```

### Fit the model

```{r}
fit_H6<-cfa(model=model_H6,
            data=dat2019.gmc,
            missing="fiml")
```


Inspect fit of the model

```{r}

round(inspect(fit_H6,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

The fit of the model is ok based on rmsea and srmr, but poor according to cfi and tli

Hypothesis 6

Print standardized estimates to test the difference between correlations

```{r}
std.est_H6<-standardizedsolution(fit_H6)
std.est_H6[std.est_H6$op==":=" | 
               std.est_H6$op=="~~" & 
               std.est_H6$lhs!=std.est_H6$rhs,]

#save to a file
export(std.est_H6[std.est_H6$op!="~1",c(1:8)],
           "../results/Table4_Unconfounded_H6_standardized_estimates.xlsx",
       overwrite=T)
```

H6. The correlation between VAA_LR and CS Self-placement on LR is strong (.469, p < .001) and stronger than the association between VAA_LR and placement of imagined party voter (.069, p = .250; difference .400, p < .001)

### Exploratory analysis of H6: Look for misspecifications


Residual correlations

```{r}
mis.rescor_H6<-miPowerFit(fit_H6,cor=.20)
mis.rescor_H6<-mis.rescor_H6[mis.rescor_H6$op=="~~" & 
                                   mis.rescor_H6$lhs!=mis.rescor_H6$rhs,]
#see summary of the decisions
table(mis.rescor_H6$decision.pow)

#there are is a single misspecification with .15 as criterion

rounded.vars<-c("mi","epc","target.epc",
                "std.epc","se.epc")

num.round<-function(var){
  var<-as.numeric(var)
  var<-round(var,2)
  return(var)
}

mis.rescor_H6[,rounded.vars]<-sapply(mis.rescor_H6[,rounded.vars],num.round)

printed.vars<-c("lhs","op","rhs","mi","epc","target.epc",
                "std.epc","std.target.epc","significant.mi",
                "high.power","decision.pow","se.epc")

#print the output

mis.rescor_H6 %>%
  filter(mis.rescor_H6$decision.pow=="M" | 
                mis.rescor_H6$decision.pow=="EPC:M") %>%
  dplyr::select(all_of(printed.vars)) 

  
```

Same misspecifications as there were for model for H3 and H4

Add to the model

```{r}
model_H6.re<-paste0(model_H6,
                      "h25~~y19\n",
                    "C2a~~C2f")

```

### Fit the respecified model

```{r}
fit_H6.re<-cfa(model=model_H6.re,
              data=dat2019.gmc,
              missing="fiml")
```

### Results

```{r}
round(inspect(fit_H6,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)

round(inspect(fit_H6.re,"fit")
      [c("npar","df","chisq","pvalue","cfi","tli","rmsea","srmr")],3)
```

Fit is improved.


Print standardized estimates to test the difference between correlations

```{r}
std.est_H6.re<-standardizedsolution(fit_H6.re)
std.est_H6.re[std.est_H6.re$op==":=" | 
               std.est_H6.re$op=="~~" & 
               std.est_H6.re$lhs!=std.est_H6.re$rhs,]

#save to a file
export(std.est_H6.re[std.est_H6.re$op!="~1",1:8],
           "../results/TableS2_Unconfounded_H6.exp_standardized_estimates.xlsx",overwrite=T)

```

Results are virtually identical.

H6. The correlation between VAA_LR and CS Self-placement on LR is moderately strong (.494, p < .001) and stronger than the association between VAA_LR and placement of imagined party voter (.072, p = .247; difference .422, p < .001)



\newpage

# Session information

```{r}
s<-sessionInfo()
print(s,locale=F)
```